---
title: "Mixed Effects Models"
author: "Ian Moran"
date: "November 3, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Mixed Effects Models
This is like previous BMD work where we model a bunch of curves separately then combine them but in this case we model all at the same time.
Useful for capturing variation across days when repeating experiments across days
mixed models included fixed factors and factors that are random variables
```{r}
library(drc)
library(drcData)
library(devtools)
install_github("DoseResponse/medrc")

```
```{r}
str(vinclozolin)
```

###Naive approach, ignore the fact that the 6 experiments are not true replicates (pseduoreplication)
```{r}
vinclo.nonlin<-drm(effect~conc, data=vinclozolin,fct=LL.4())
summary(vinclo.nonlin)
```

##Fitting mixed effect model
```{r}
library(medrc)

vinclo.nlmm<-medrm(effect~conc,data=vinclozolin, fct=LL.4(),
                   random= d~1|exper)
```
This is simplest way to specify.. says that parameter d, the 
alternatively could assume e+d~1|exper

```{r}
summary(vinclo.nlmm)
```

negative c is not great, maybe limit to 3 parameter model, not today

Compare EC values across naive and mixed model
```{r}
ED(vinclo.nonlin, c(10,20,50))

ED(vinclo.nlmm, c(10,20,50))
```

##Last example Species sensitivity distributions
















